Ecological management of stochastic systems with long transients

### 2019-12-31 manuscript/manuscript.Rmd
WIP

```{r fig_model, fig.cap="Change in population size $X_{t+1} - X_t = f(X_t)$. Equilibria occur when the change curve crosses zero, stable points arise at crossings with negative slopes, unstable equilibria have positive slopes.  By changing parameter $a$, the system goes from bistable (red, $a = 28$) to a ghost attractor (blue, $a = 27.2$). For both curves, $r = 0.8$, $K = 153$, $Q = 2$, $H = 20$."}

@@

```{r fig_sims}
# Fig 2
fig1_sims %>%
  ggplot(aes(time, state, group = interaction(rep, scenario), col= scenario)) + 
  geom_line(alpha=0.3) + scale_color_solarized() + facet_wrap(~scenario)
```

@@

```{r fig_policy, fig.cap="Optimal policy functions under each model. Parameter values for utility function used here are $c = 1$, $d = 0.5$, $X_0 = 50$, $\delta = 0.999$, parameters for the transition models are as given in Fig 1. "}
## Fig 3

@@

```{r fig_sims_by_prior}
# Fig 4
fixed_sims <- read_csv("manuscript_cache/fixed_sims.csv")

@@

```{r}
costs <- fixed_sims %>% group_by(time, prior)  %>% 
  summarise(cost = mean(value)) %>%
  group_by(prior) %>%
  mutate(cost = abs(cumsum(cost * .999 ^ time))) %>%
  rename(model = prior)

@@

```{r fig_sims_by_prior}
## Fig 6

@@

```{r mdp_learning}
# Fig 7
set.seed(12345)
Tmax <- 50
learning <- mdp_learning(list(P_ghost, P_bistable), reward, discount, 
                   model_prior = c(0.5, 0.5),
                   x0 = x0, Tmax = Tmax,
                  true_transition = P_ghost,
                  max_iter = 1e4, epsilon = 1e-2)

@@

```{r fig_learning_sim}
# Fig 8
 learning$df %>% 
  select(-value) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, states[state], color = series)) + geom_line()
```

```{r fig_learning_posterior}
model_names <- c("Ghost", "Bistable")
learning$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% floor(seq(1,10, length.out=5))) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = model_names[as.integer(param)]) %>% 
  ggplot(aes(x = param, y = probability, group = time, alpha = time)) + 
  geom_bar(stat="identity", position = "identity", show.legend = FALSE, fill=pal[1])
```

@@

- Long transients can be misleading -- e.g. mistaking a ghost attractor for a stable state.
- These mistakes can be significant, since long transients require qualitatively different management strategies.
- Nevertheless, thes

>>>>

```{r figure1, fig.cap="Change in population size $X_{t+1} - X_t = f(X_t)$. Equilibria occur when the change curve crosses zero, stable points arise at crossings with negative slopes, unstable equilibria have positive slopes.  By changing parameter $a$, the system goes from bistable (red, $a = 28$) to a ghost attractor (blue, $a = 27.2$). For both curves, $r = 0.8$, $K = 153$, $Q = 2$, $H = 20$."}

@@

```{r figure2}
# Fig 2
fig1_sims %>%
  ggplot(aes(time, state, group = interaction(rep, scenario), col= scenario)) + 
  geom_line(alpha=0.3) + scale_color_solarized() + facet_wrap(~scenario)
```

@@

```{r figure3, fig.cap="Optimal policy functions under each model. Parameter values for utility function used here are $c = 1$, $d = 0.5$, $X_0 = 50$, $\\delta = 0.999$, parameters for the transition models are as given in Fig 1."}
## Fig 3

@@

```{r figure4}
# Fig 4
fixed_sims <- read_csv("manuscript_cache/fixed_sims.csv")

@@

```{r figure5}
costs <- fixed_sims %>% group_by(time, prior)  %>% 
  summarise(cost = mean(value)) %>%
  group_by(prior) %>%
  mutate(cost = abs(cumsum(cost * .999 ^ time))) %>%
  rename(model = prior)

@@

```{r figure6}
## Fig 6

@@

```{r mdp_learning}
set.seed(12345)
Tmax <- 50
learning <- mdp_learning(list(P_ghost, P_bistable), reward, discount, 
                   model_prior = c(0.5, 0.5),
                   x0 = x0, Tmax = Tmax,
                  true_transition = P_ghost,
                  max_iter = 1e4, epsilon = 1e-2)

@@

```{r figure7}
# Fig 7
 learning$df %>% 
  select(-value) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, states[state], color = series)) + geom_line()
```

```{r figure8}
model_names <- c("Ghost", "Bistable")
learning$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% floor(seq(1,10, length.out=5))) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = model_names[as.integer(param)]) %>% 
  ggplot(aes(x = param, y = probability, group = time, alpha = time)) + 
  geom_bar(stat="identity", position = "identity", show.legend = FALSE, fill=pal[1])
```

@@

- Long transients can be misleading -- e.g. mistaking a ghost attractor for a stable state.
- These mistakes can be significant, since long transients require qualitatively different management strategies.
- Decision-theoretic approaches provide a way to manage this uncertainty

Markov Decision Processes are an active area of research in robotics and artificial intelligence community, though terminology differs substantially.
